Risk and Trust in Open Systems:\\ \Large Towards formalising Permission and Authority, and specifying policies for Object Capability Patterns

### 2017-09-07 shupeng_msc_report.tex
update

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
\newcommand{\reporttitle}{Risk and Trust in Open Systems}
\newcommand{\reportauthor}{Shu Peng Loh}
\newcommand{\supervisor}{Sophia Drossopoulou}
\newcommand{\degreetype}{Computing Science}
\documentclass[a4paper,11pt,twoside]{article}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[hang,flushmargin]{footmisc} 
\definecolor{light-gray}{gray}{0.95}
\usepackage{listings}
\lstset{
    breaklines=true,
    columns=flexible,
    backgroundcolor=\color{light-gray},
    basicstyle=\sffamily\small,
    escapechar=\&
}
\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.15} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics
\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage{multirow}
\usepackage{multicol}
\usepackage[left=2.5cm,right=2.5cm,top=3cm,bottom=3cm, hmarginratio=1:1,headheight=14pt,columnsep=15pt]{geometry} % Document margins
\usepackage[hang, normal,labelfont=bf,up,textfont=sc,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{enumitem} % Customized lists
\usepackage{color}
\usepackage{pdfpages}
\setlist[itemize]{noitemsep} % Make itemize lists more compact
\usepackage{ latexsym }
\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small} % Set the abstract itself to small italic text

@@

\makeatletter
\DeclareRobustCommand{\emp}{%
  \@nomath\em \if b\expandafter\@car\f@series\@nil
  \normalfont \else \sffamily \bfseries \fi}
\makeatother
\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent  We revisit the concepts of \textit{permission} and \textit{authority} in the works of Miller\cite{miller2006} and \linebreak Drossopoulou et al. \cite{drossopoulou2016}, and we contribute to the literature by proposing new formal definitions of permission and authority (MayAccess and MayCall predicates in our paper). We show how our definitions can be used to reason about object-to-object interactions, and in particular, cooperation between objects and protection of objects from unknown code. Another contribution we made is that we propose and show how the concept of domination over objects can be used to specify and reason about safe or vulnerable cooperation, which is novel but inspired by the work of Clarke et al. on ownership types \cite{clarke1998}. Our paper then show how using a mix of our proposed formalisations of permission and authority, and Hoare logics, we can formally specify well-established Object-Capability (OCap) patterns in the form of \textit{OCap policies} \cite{drossopoulou2013,drossopoulou2014,drossopoulou2015,drossopoulou2015b}.  In particular, we reason about protecting the properties of nodes in a DOM Tree, inspired by the works of Maffeis et al.\cite{maffeis2010} and Devriese et al.\cite{devriese2016}, but we do so within the capability-safe language Pony \cite{clebsch2015}. We believe our methodology and specifications of OCap patterns are less complex than other methodologies in the literature\cite{devriese2016,swasey2017}, which will allow a programmer to reason convincingly how the use of attenuating objects in capability-safe languages can help to facilitate cooperation betweeen known and unknown code, while still preventing vulnerabilities by preserving desired properties in an open system. Lastly, we end our paper with some preliminary insights on thinking about security and risk in a non-OCap system by using the Ethereum\cite{wood2014} blockchain as a case study.\end{abstract}
}
\begin{document}
\let\sf\textsf
\input{titlepage}
\raggedbottom
\thispagestyle{empty}
\tableofcontents
\thispagestyle{empty}

\maketitle
\setcounter{page}{1}
\pagestyle{fancy} % All pages have headers and footers

@@

With our formal definitions of permission, authority, and domination, our paper then argues how we can use them to reason and distinguish concepts of \textit{isolation}, \textit{cooperation}, \textit{vulnerability} and \textit{protection}. We claim that our framework allows one to understand and reason about these concepts in a highly straightforward and intuitive way without treating security as a separate burden from the functionality of the system, which we hope would encourage programmers to seek the use of the object-capability model to build robust and secure systems.\\

To show this in practice, our paper provides an illustration of three OCap patterns found in \cref{sec:figdom,sec:figcaretaker,sec:figmem}, but we focus only on providing the OCap policy specifications for the DOM Tree pattern in \cref{sec:dompolicies}. We choose the DOM Tree pattern because we are convinced of the importance of reasoning about unknown code using the object-capability framework in the web security space, where guest/unknown code (such as third party scripted advertising or widgets) are pervasive in the web today (details can be found in the work by Maffeis et al.\cite{maffeis2010}). We show how our policies can specify the preservation of properties of nodes, in the presence of unknown code execution. Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain system that has experienced explosive growth in recent years, and we recognise that further work has to be done in this exciting new field.

\subsection{Overview of Paper}
In our paper, we first give an overview background history on the OCap model in \cref{sec:ocapmodel}. Readers familiar with the object-capability model and its languages are encouraged to skim through this section. In \cref{sec:specs}, we seek to formally specify the definitions of permission and authority and use our specifications to describe and reason how objects interact with each other. Our formal specifications are broad enough to reason in general about how two objects interact in our small specification language. That said, there are benefits to using these concepts to reason within an OCap system, which we demonstrate in \cref{sec:ocapspecs}, because the restrictions placed by OCap languages or what we call capability-safe languages allow us to enforce protection of objects more easily. We further solidify how our methodology can be used to reason convincgly in an OCap system by applying it on a DOM Tree pattern within an OCap framework, in \cref{sec:ocappatterns}.
%----------------------------------------------------------------------------------------
%   OCap Model
%----------------------------------------------------------------------------------------
\section{Object-Capability Model}\label{sec:ocapmodel}
\subsection{From Object to Capability}
The OCap model uses the reference graph of the objects as the access graph, and strictly requires objects to interact with each other only by sending messages on object references\cite{miller2003b}. When these references cannot be forged in the system, and when there there are no default globally accessible objects (for an object to be globally accessible, the reference to use that object has to be \textit{explicitly} granted to every object globally), then it becomes necessary that for an object to call the methods of another object, an object needs to explicitly obtain the reference of the object it is calling.  Object references therefore embody authority, and we use the word capability to describe the encapsulation of both reference (designation) and authority which are unforgeable.

@@

\subsection{Capability-safe Language Restrictions}
In this section, we briefly mention three critical restrictions required in a capability-safe language: memory safety, no global mutable variables (objects can only interact with each other on the reference graph), and the taming of reflection.
\subsubsection{No Capability Forging/Memory-safety}
The object-capability model requires a capability to be \textit{unforgeable}. Without memory safety in the programming language, a reference to a particular object in a memory can potentially be forged externally and anywhere. For example in C++, because the language allows memory manipulation, designation is not opaque and hence forgeable. Let us assume a pointer ptr that holds the designation of a newly created int object \texttt{\small int* ptr = new int()}. The int object can thus be accessed  through the use of ptr. However C++ allows the programmer access to the memory and we can obtain the a string of bits that represent the memory address of the int object. Let us assume this memory address is represented by the bits 0x12345678. We say Designation is forgeable and accessible by bits because any object can perform the following operation:\\
\ablock\texttt{\small int* ptr2 = reinterpret\_cast<int*>(0x12345678)}\\
to obtain a new pointer variable to the same int object. On the other hand, in java, object designations are completely opaque/indivisible--even with knowledge of the bits that constitute a pointer (obtained by Object.toString()); there is no operation for obtaining the corresponding pointer from this bit information and therefore designation cannot be forged.
\subsubsection{No Global Mutable Variables}
This restriction is necessary because if there are global mutable variables, then objects can now communicate without needing the object capabilities of each other. Global mutable variables can behave like a message carrier, such that two objects can both communicate by monitoring and modifying the state of the global variable. The capability-safe language Pony, which we use to show the code of OCap patterns, do not allow global variables.
\subsubsection{No 'Untamed' Reflection}
Reflection is the ability of a piece of code to "reflect" on the structure of other code in the same system or itself. More concretely in object-oriented systems, an object that can perform untamed reflection is about to examine and modify the code of other objects or itself. Take for example in Java, reflection allows code to perform operations that would otherwise be illegal in non-reflective code, such as accessing private fields and methods. Let us assume a simple Java program where object A possesses the reference capability of object B, and object B possesses the reference capability of object C. Object B stores the reference capability of object C in a private field and does not have any public method to divulge this data. With reflection, we cannot guarantee that object A will never obtain the capability of object C - this is because with the capability of object B, object A can reflect on object B's code and get all information stored in object B - including the private field that holds the reference capability of object C. In order to ensure capability-safety, the reflection feature in a language has to be 'tamed' or restricted. Joe-E\cite{mettler2010}, a security oriented subset of Java, has such a feature.
\subsection{Related Works}

@@

%----------------------------------------------------------------------------------------
%   Formal Specifications
%----------------------------------------------------------------------------------------
\newpage
\section{Formal specifications}\label{sec:specs}
In this section we describe a small object-oriented language in \cref{sec:speclang} and define predicates that allow us to specify object-capability policies in \cref{sec:permauth}. For our language, we use the definitions of modules, classes, runtime state, interpretation, execution and arising configurations from the specification language in the appendix of \cite{drossopoulou2015b}. To express more precisely how an object can interact with another object, we use the ideas of permission and authority in Miller's\cite{miller2006} and Drossopoulou et al.'s\cite{drossopoulou2016} works. In their papers and our paper, we generally try to capture the informal meaning of:
\begin{itemize}
\item permission: access right of an object to invoke a behaviour of an object
\item authority: ability to cause effects
\end{itemize}
In the approach of our paper however, we contribute to the literature by taking a novel approach in formally defining both permission and authority, though we argue they generally capture the same informal meanings. For permission, we have four different modes permissions specified by four MayAccess predicates. Given a module M and state \losigma,:
\begin{tabbing}
MayAccess$^{Dir,Now}$(o,o$'$): \=o has an object reference of o$'$\\
MayAccess$^{Ind,Now}$(o,o$'$): o contains a reference that points to an object that has a reference\\
\>to another object---thus forming a series of object reference(s)---\\ \> where the last object in the series has an object reference of o$'$\\
MayAccess$^{Dir,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(o,o$'$) holds\\
MayAccess$^{Ind,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(o,o$'$) holds\\
\end{tabbing}
For authority, we propose a new MayCall predicate that deviates from the MayAffect predicate proposed in Drossopoulou et al.\cite{drossopoulou2016}. While their MayAffect predicate captures the idea of a field of an object being modified, our MayCall predicate is much weaker in the sense that mean the ability to invoke a method of an object \textit{being}. Given a module M and state \losigma,:
\begin{tabbing}
MayCall(o,o$'$): \=there exist a method in o that can call a method in o$'$,\\ \> such that there will be some eventual state \losigma$'$ arising from \losigma,\\ \> where the receiver in the stack has transitioned from o in \losigma, to o$'$ in \losigma$'$\\
\end{tabbing}

@@

where, we say MayAccess$^{Dir,Now}$(x,e) holds \lotiff\ in some current state \losigma, we have a field in x that points to e, or that we have x as a receiver in \losigma\ and there exists some y in the same \losigma\ that points to e. We also say MayAccess$^{Ind,Now}$(x,e) holds \lotiff\ in some current state \losigma, we have a \textit{series of fields} from x that leads to e, or that we have x as a receiver in \losigma\ and there exists some y with a series of fields in the same \losigma\ that leads to e. Lastly, the 'eventual' definitions of the two predicates are then defined as there existing an arising configuration \losigma$'$ from \losigma, where the 'now' definitions hold at \losigma$'$.\\

Note that MayAccess$^{\_,Now}$ holds imply MayAccess$^{\_,Eve}$ holds, since an eventual state \losigma$'$ that arises from \losigma\, can refer to \losigma, and therefore, the 'now' definitions implies and are stronger than the 'eventual' definitions. Also since a series of fields \loconj{f}\ can mean a singular field f, MayAccess$^{Dir,\_}$ holds imply MayAccess$^{Ind,\_}$ holds, and therefore the 'direct' definitions are stronger than the 'indirect' definitions. \\
\begin{table}[htb]
\caption{MayAccess Relations}\label{tab:accessrelations}
\centering
\begin{tabular*}{0.5\linewidth}{c|ccc}\toprule
& \bf Now & & \bf Eventually\\
\hline
\multirow{5}{*}{\rot{\bf Indirect \enspace Direct \:}} & \multirow{2}{*}{MayAccess$^{Dir,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Dir, Eve}$} \\
& & \lonimpliedby &  \\
& \rot{\loimpliedby} \rot{\lonimplies}& &\rot{\loimpliedby} \rot{\lonimplies} \\
& \multirow{2}{*}{MayAccess$^{Ind,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Ind, Eve}$} \\
& & \lonimpliedby &\\
\end{tabular*}
\end{table}\\
We surmarise the relationships between the four flavours of MayAccess in \cref{tab:accessrelations}. Note that in an object-to-object paradigm, all four flavours of MayAccess involve object capabilities. MayAccess$^{Dir,\_}$(o,o$'$) says object o has the direct reference of o$'$, while MayAccess$^{Ind,\_}$ says object o has a path of reference(s) that lead to o$'$ (that is, o has the reference of some object x$_1$, and object x$_1$ has the reference of object x$_2$ $\dots$ and object x$_n$ has the reference of object o$'$). The 'now' and 'eventual' flavours simply state that the "direct" and "indirect" predicates either they hold at \losigma\ (Now) or they can hold at some arising eventual state \losigma' from \losigma\ (eventual).

\subsubsection{Authority - MayCall Definition}\label{MayCall}
With our access predicates in place, we define a MayCall predicate that describe what it means for x to be able to call e. Note that our definition of MayCall is different from the MayAffect predicate in \cite{drossopoulou2016}. Our MayCall predicate captures the meaning of an object o having authority over another object o$'$ without o necessarily modifying the fields of o$'$. In that sense, our MayCall predicate is weaker than the MayAffect predicate in \cite{drossopoulou2016}:\\
\begin{logic}
\hr\\
\emp{Definition---[MayCall]}\\
M,\losigma\ \loturns\ \=MayCall(x,e) \loiff \\
\> \loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists \losigma$'$\loin Arising(M,\losigma). \loand\ \loexec{\texttt{this}}{\losigma$'$} = \loexec{e}{\losigma$'$}\\
\hr
\end{logic}

@@

\subsection{Domination}\label{sec:domination}
Here, we introduce the concept of domination that is inspired by the work on ownership types by Clarke et al.\cite{clarke1998}, where they define that the owner o of an object o$'$ dominates o$'$ in the object graph such that for any path p from some arbitrary object o$'$' to o$'$, either o$'$' belongs to the set of owners or that p passes through o.\\

We say that a set S dominates an object x \lotiff\ for every object y in the system that has a path to x, y must have a path to some arbitrary object o that is a member of the set S. The power of having such a predicate will be demonstrated in \cref{sec:ocappatterns}, where we show how attenuating objects that is meant to protect an object o can ensure certain properties of members in a dominating set over s under any unknown code execution.\\

\begin{logic}
\hr\\
\emp{Definition---[Domination]}\\
M,\losigma\ \loturns\ \=Dom(S,x) \loiff \\
\> \loforall y.\loconj{f},n. [ \= x\loneq y \loand\ \loexec{y.f$_1$...f$_n$}{\losigma} = \loexec{x}{\losigma} \loimplies\\
\>\> \loexists k,o. \loexec{y.f$_1$...f$_k$}{\losigma} = \loexec{o}{\losigma} \loand\ o\loin S ]\\
\hr
\end{logic}
The usefulness of domination will be made more clear in the next section when we argue about cooperation without vulnerability.
\section{Reasoning in an OCap System}\label{sec:ocapspecs}

In this section, we formally describe the ideas of \textit{isolation} and \textit{cooperation}, and how \textit{cooperation} can either be vulnerable or protected, within the OCap system. We use the word \textit{capability} (of an object), interchangeably with reference (of an object), to represent that both object reference and the right to use the object reference are indistinguishable in OCap; i.e., there is no global ambient authority to prevent an object from using another object reference in general. Given our definition of \textit{permission}, it follows that o has the permission of o$'$ if o has the capability of o'. However, if we were to use \textit{our} formal definition of \textit{authority} (causing effects where effects are method invocations), we cannot say that an object reference embodies authority - we elaborate more in \cref{sec:isolation and cooperation}. We use the word \textit{path} interchangeably with indirect capability or permission, to emphasise that an object o need not necessarily require the direct capability of o$'$ to invoke the behaviours of o$'$---invocation of o$'$ can be done by o through a network of paths consisting of other object capabilities.\\

>>>>

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
\newcommand{\reporttitle}{Risk and Trust in Open Systems}
\newcommand{\reportauthor}{Shu Peng Loh}
\newcommand{\supervisor}{Sophia Drossopoulou}
\newcommand{\degreetype}{Computing Science}
\documentclass[a4paper,11pt,twoside]{article}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[hang,flushmargin]{footmisc} 
\definecolor{light-gray}{gray}{0.95}
\usepackage{listings}
\lstset{
    breaklines=true,
    columns=flexible,
    backgroundcolor=\color{light-gray},
    basicstyle=\sffamily\small,
    escapechar=\&
}
\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.15} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics
\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage{multirow}
\usepackage{multicol}
\usepackage[left=2.5cm,right=2.5cm,top=3cm,bottom=3cm, hmarginratio=1:1,headheight=14pt,columnsep=15pt]{geometry} % Document margins
\usepackage[hang, normal,labelfont=bf,up,textfont=sc,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage{enumitem} % Customized lists
\usepackage{color}
\usepackage{pdfpages}
\setlist[itemize]{noitemsep} % Make itemize lists more compact
\usepackage{ latexsym }
\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small} % Set the abstract itself to small italic text

@@

\makeatletter
\DeclareRobustCommand{\emp}{%
  \@nomath\em \if b\expandafter\@car\f@series\@nil
  \normalfont \else \sffamily \bfseries \fi}
\makeatother
\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent  We revisit the concepts of \textit{permission} and \textit{authority} in the works of Miller\cite{miller2006} and \linebreak Drossopoulou et al. \cite{drossopoulou2016}, and we contribute to the literature by proposing new formal definitions of permission and authority (MayAccess and MayCall predicates in our paper). We show how our definitions can be used to reason about object-to-object interactions, and in particular, cooperation between objects and protection of objects from unknown code. Another contribution we made is that we propose and show how the concept of domination over objects can be used to specify and reason about safe or vulnerable cooperation, which is novel but inspired by the work of Clarke et al. on ownership types \cite{clarke1998}. Our paper then show how using a mix of our proposed formalisations of permission, authority, domination, and Hoare logics, we can formally specify well-established Object-Capability (OCap) patterns in the form of \textit{OCap policies} \cite{drossopoulou2013,drossopoulou2014,drossopoulou2015,drossopoulou2015b}.  In particular, we reason about protecting the properties of nodes in a DOM Tree, inspired by the works of Maffeis et al.\cite{maffeis2010} and Devriese et al.\cite{devriese2016}, but we do so within the capability-safe language Pony \cite{clebsch2015}. We believe our methodology and specifications of OCap patterns are less complex than other methodologies in the literature\cite{devriese2016,swasey2017}, which will allow a programmer to reason convincingly how the use of attenuating objects in capability-safe languages can help to facilitate cooperation betweeen known and unknown code, while still preventing vulnerabilities by preserving desired properties in an open system. Lastly, we end our paper with some preliminary insights on thinking about security and risk in a non-OCap system, using the Ethereum\cite{wood2014} blockchain as a case study.\end{abstract}
}
\begin{document}
\let\sf\textsf
\input{titlepage}
\raggedbottom
\thispagestyle{empty}
\tableofcontents
\thispagestyle{empty}
\clearpage
\thispagestyle{empty}
\begin{minipage}{\textwidth}
\begin{center}
\textit{Acknowledgements}
\end{center}
\small
I am immensely grateful to my supervisor Prof. Sophia Drossopoulou, for her dedication and guidance, where many ideas included in this thesis were derived and inspired from. I will miss the countless hours we have worked together on this thesis.\\

I am also thankful for the emotional support and encouragement from my friends and family; and the Monetary Authority of Singapore for offering me a scholarship to pursue this MSc degree at Imperial College London.
\end{minipage}
\maketitle
\setcounter{page}{1}
\pagestyle{fancy} % All pages have headers and footers

@@

With our formal definitions of permission, authority, and domination, our paper then argues in \cref{sec:ocapreasoning} on how we can use them to reason and distinguish concepts of \textit{isolation}, \textit{cooperation}, \textit{vulnerability} and \textit{protection}. We claim that our framework allows one to understand and reason about these concepts in a highly straightforward and intuitive way without treating security as a separate burden from the functionality of the system, which we hope would encourage programmers to seek the use of the object-capability model to build robust and secure systems.\\

To show this in practice, our paper provides an illustration of three OCap patterns found in \cref{sec:figdom,sec:figcaretaker,sec:figmem}, and we focus only on providing the OCap policy specifications for the DOM Tree pattern in \cref{sec:dompolicies}. We choose the DOM Tree pattern because we are convinced of the importance of reasoning about unknown code using the object-capability framework in the web security space, where guest/unknown code (such as third party scripted advertising or widgets) are pervasive in the web today (details can be found in the work by Maffeis et al.\cite{maffeis2010}). We show how using our OCap policies, we can specify the preservation of properties of nodes, in the presence of unknown code execution, in \cref{sec:treereasoning}. Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain platform that has experienced explosive growth in recent years, and we recognise that further work has to be done in the field of smart contracts on the blockchain technology.

\subsection{Overview of Paper}
In our paper, we first give an overview background history on the OCap model in \cref{sec:ocapmodel}. Readers familiar with the object-capability model and its languages are encouraged to skim through this section. In \cref{sec:specs}, we seek to formally specify the definitions of permission, authority and domination, and we use our specifications to describe and reason how objects interact with each other. Our formal specifications are broad enough to reason in general about how two objects interact in our small specification language. That said, there are benefits to using these concepts to reason within an OCap system, which we demonstrate in \cref{sec:ocapreasoning}, because the restrictions placed by OCap languages or what we call capability-safe languages allow us to enforce protection of objects more easily. We further solidify how our methodology can be used to reason convincingly in an OCap system by applying it on a DOM Tree pattern within an OCap framework, in \cref{sec:ocappatterns}. We end off with some preliminary insights on Ethereum in \cref{sec:ethereum}.
%----------------------------------------------------------------------------------------
%   OCap Model
%----------------------------------------------------------------------------------------
\section{Object-Capability Model}\label{sec:ocapmodel}
\subsection{From Object to Capability}
The OCap model uses the reference graph of the objects as the access graph, and strictly requires objects to interact with each other only by sending messages on object references\cite{miller2003b}. When these references cannot be forged in the system, and when there there are no default globally accessible objects (for an object to be globally accessible, the reference to use that object has to be \textit{explicitly} granted to every object globally), then it becomes necessary that for an object to call the methods of another object, an object needs to explicitly obtain the reference of the object it is calling.

@@

\subsection{Capability-safe Language Restrictions}
In this section, we briefly mention three critical restrictions required in a capability-safe language: memory safety, no global mutable variables (objects can only interact with each other on the reference graph), and the taming of reflection.
\subsubsection{No Capability Forging/Memory-safety}
The object-capability model requires a capability to be \textit{unforgeable}. Without memory safety in the programming language, a reference to a particular object in a memory can potentially be forged externally and anywhere. For example in C++, because the language allows memory manipulation, designation is not opaque and hence forgeable. Let us assume a pointer ptr that holds the designation of a newly created int object \texttt{\small int* ptr = new int()}. The int object can thus be accessed  through the use of ptr. However C++ allows the programmer access to the memory and we can obtain the a string of bits that represent the memory address of the int object. Let us assume this memory address is represented by the bits 0x12345678. We say Designation is forgeable and accessible by bits because any object can perform the following operation:\\
\ablock\texttt{\small int* ptr2 = reinterpret\_cast<int*>(0x12345678)}\\
to obtain a new pointer variable to the same int object. On the other hand, in java, object designations are completely opaque/indivisible--even with knowledge of the bits that constitute a pointer (obtained by Object.toString()); there is no operation for obtaining the corresponding pointer from this bit information and therefore designation cannot be forged.
\subsubsection{No Global Mutable Variables}
This restriction is necessary because if there are global mutable variables, then objects can now communicate without needing the object capabilities of each other. Global mutable variables can behave like a message carrier, such that two objects can both communicate by monitoring and modifying the state of the global variable. The capability-safe language Pony\cite{clebsch2015}, which we use to show the code of OCap patterns, do not allow global variables.
\subsubsection{No 'Untamed' Reflection}
Reflection is the ability of a piece of code to "reflect" on the structure of other code in the same system or itself. More concretely in object-oriented systems, an object that can perform untamed reflection is about to examine and modify the code of other objects or itself. Take for example in Java, reflection allows code to perform operations that would otherwise be illegal in non-reflective code, such as accessing private fields and methods. Let us assume a simple Java program where object A possesses the reference capability of object B, and object B possesses the reference capability of object C. Object B stores the reference capability of object C in a private field and does not have any public method to divulge this data. With reflection, we cannot guarantee that object A will never obtain the capability of object C - this is because with the capability of object B, object A can reflect on object B's code and get all information stored in object B - including the private field that holds the reference capability of object C. In order to ensure capability-safety, the reflection feature in a language has to be 'tamed' or restricted. Joe-E\cite{mettler2010}, a security oriented subset of Java, has such a feature.
\subsection{Related Works}

@@

%----------------------------------------------------------------------------------------
%   Formal Specifications
%----------------------------------------------------------------------------------------
\newpage
\section{Formal specifications}\label{sec:specs}
In this section we describe a small object-oriented language in \cref{sec:speclang} and define predicates that allow us to specify object-capability policies in \cref{sec:permauth}. For our language, we use the definitions of modules, classes, runtime state, interpretation, execution and arising configurations from the specification language in the appendix of \cite{drossopoulou2015b}. To express more precisely how an object can interact with another object, we use the ideas of permission and authority in Miller's\cite{miller2006} and Drossopoulou et al.'s\cite{drossopoulou2016} works. In their papers and our paper, we generally try to capture the informal meaning of:
\begin{itemize}
\item permission: access right of an object to invoke a behaviour of an object
\item authority: ability to cause effects
\end{itemize}
In the approach of our paper however, we contribute to the literature by taking a novel approach in formally defining both permission and authority, though we argue they generally capture the same informal meanings. For permission, we have four different modes permissions specified by four MayAccess predicates. Given a module M and state \losigma,:
\begin{tabbing}
MayAccess$^{Dir,Now}$(o,o$'$): \=o has an object reference of o$'$\\
MayAccess$^{Ind,Now}$(o,o$'$): o contains a reference that points to an object that has a reference\\
\>to another object---thus forming a series of object reference(s)---\\ \> where the last object in the series has an object reference of o$'$\\
MayAccess$^{Dir,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(o,o$'$) holds\\
MayAccess$^{Ind,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(o,o$'$) holds\\
\end{tabbing}
For authority, we propose a new MayCall predicate that deviates from the MayAffect predicate proposed in Drossopoulou et al.\cite{drossopoulou2016}. While their MayAffect predicate captures the idea of a field of an object being modified, our MayCall predicate is much weaker ---we mean the ability to invoke a method of an object. Therefore implicitly, we consider method invocation as an effect in the system. Given a module M and state \losigma,:
\begin{tabbing}
MayCall(o,o$'$): \=there exist a method in o that can call a method in o$'$,\\ \> such that there will be some eventual state \losigma$'$ arising from \losigma,\\ \> where the receiver in the stack has transitioned from o in \losigma, to o$'$ in \losigma$'$\\
\end{tabbing}

@@

where, we say MayAccess$^{Dir,Now}$(x,e) holds \lotiff\ in some current state \losigma, we have a field in x that points to e, or that we have x as a receiver in \losigma\ and there exists some y in the same \losigma\ that points to e. We also say MayAccess$^{Ind,Now}$(x,e) holds \lotiff\ in some current state \losigma, we have a \textit{series of fields} from x that leads to e, or that we have x as a receiver in \losigma\ and there exists some y that leads to e. Lastly, the 'eventual' definitions of the two predicates are then defined as there existing an arising configuration \losigma$'$ from \losigma, where the 'now' definitions hold at \losigma$'$.\\

Note that MayAccess$^{\_,Now}$ holds imply MayAccess$^{\_,Eve}$ holds, since an eventual state \losigma$'$ that arises from \losigma\, can refer to \losigma, and therefore, the 'now' definitions implies and are stronger than the 'eventual' definitions. Also since a series of fields \loconj{f}\ can mean a singular field f, MayAccess$^{Dir,\_}$ holds imply MayAccess$^{Ind,\_}$ holds, and therefore the 'direct' definitions are stronger than the 'indirect' definitions. \\
\begin{table}[htb]
\caption{MayAccess Relations}\label{tab:accessrelations}
\centering
\begin{tabular*}{0.5\linewidth}{c|ccc}\toprule
& \bf Now & & \bf Eventually\\
\hline
\multirow{5}{*}{\rot{\bf Indirect \enspace Direct \:}} & \multirow{2}{*}{MayAccess$^{Dir,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Dir, Eve}$} \\
& & \lonimpliedby &  \\
& \rot{\loimpliedby} \rot{\lonimplies}& &\rot{\loimpliedby} \rot{\lonimplies} \\
& \multirow{2}{*}{MayAccess$^{Ind,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Ind, Eve}$} \\
& & \lonimpliedby &\\
\end{tabular*}
\end{table}\\
We surmarise the relationships between the four flavours of MayAccess in \cref{tab:accessrelations}. Note that in an object-to-object paradigm, all four flavours of MayAccess involve object capabilities. MayAccess$^{Dir,\_}$(o,o$'$) says object o has the direct reference of o$'$, while MayAccess$^{Ind,\_}$ says object o has a path of reference(s) that lead to o$'$ (that is, o has the reference of some object x$_1$, and object x$_1$ has the reference of object x$_2$ $\dots$ and object x$_n$ has the reference of object o$'$). The 'now' and 'eventual' flavours simply state that the "direct" and "indirect" predicates either they hold at \losigma\ (now) or they can hold at some arising eventual state \losigma' from \losigma\ (eventual).

\subsubsection{Authority - MayCall Definition}\label{MayCall}
With our MayAccess predicates in place, we define a MayCall predicate that describe what it means for x to be able to call e. Note that our definition of MayCall is different from the MayAffect predicate in \cite{drossopoulou2016}. Our MayCall predicate captures the meaning of an object o having authority over another object o$'$ without o necessarily modifying the fields of o$'$. In that sense, our MayCall predicate is weaker than the MayAffect predicate in \cite{drossopoulou2016}:\\
\begin{logic}
\hr\\
\emp{Definition---[MayCall]}\\
M,\losigma\ \loturns\ \=MayCall(x,e) \loiff \\
\> \loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists \losigma$'$\loin Arising(M,\losigma). \loand\ \loexec{\texttt{this}}{\losigma$'$} = \loexec{e}{\losigma$'$}\\
\hr
\end{logic}

@@

\subsection{Domination}\label{sec:domination}
Here, we introduce the concept of domination that is inspired by the work on ownership types by Clarke et al.\cite{clarke1998}, where they define that the owner o of an object o$'$, dominates o$'$ in the object graph such that for any path p from some arbitrary object o$''$ to o$'$, either o$''$ belongs to the set of owners or that p passes through o.\\

We say that a set S dominates an object x \lotiff\ for every object y in the system that has a path to x, y must have a path to some arbitrary object o that is a member of the set S.

\begin{logic}
\hr\\
\emp{Definition---[Domination]}\\
M,\losigma\ \loturns\ \=Dom(S,x) \loiff \\
\> \loforall y.\loconj{f},n. [ \= x\loneq y \loand\ \loexec{y.f$_1$...f$_n$}{\losigma} = \loexec{x}{\losigma} \loimplies\\
\>\> \loexists k,o. \loexec{y.f$_1$...f$_k$}{\losigma} = \loexec{o}{\losigma} \loand\ o\loin S ]\\
\hr
\end{logic}
The usefulness of domination will be made more clear in the next section where we argue about cooperation without vulnerability. The power of having such a predicate will also be demonstrated in \cref{sec:ocappatterns}, where we show how attenuating objects that is meant to protect an object o can ensure certain properties of the system under unknown code execution.\\
\section{Reasoning in an OCap System}\label{sec:ocapreasoning}

In this section, we formally describe the ideas of \textit{isolation} and \textit{cooperation}, and how \textit{cooperation} can either be vulnerable or protected, within the OCap system. We use the word \textit{capability} (of an object), interchangeably with reference (of an object), to represent that both object reference and the right to use the object reference are indistinguishable in OCap; i.e., there is no global ambient authority to prevent an object from using another object reference. Given our 4 formal definitions of \textit{permission}, it follows that o has different kinds of permission of o$'$ depending on whether o has a direct or indirect capability of o$'$, and whether we are arguing about a current or eventual state of the system. However, if we were to use \textit{our} formal definition of \textit{authority} (causing effects where effects are method invocations), we cannot say that an object reference embodies authority - we elaborate more in \cref{sec:isolation and cooperation}. We use the word \textit{path} interchangeably with indirect capability or permission, to emphasise that an object o need not necessarily require the direct capability of o$'$ to invoke the behaviours of o$'$---invocation of o$'$ can be done by o through a network of paths consisting of other object capabilities.\\
